---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme_files/README-",
  out.width = "100%"
)
library(es4all)
```

# Effect sizes for all! (`es4all`)

<!-- badges: start -->
<!-- badges: end -->

`es4all` computes dominance-based effect sizes for all data types: continuous, binary, and ordinal data. This family of effect sizes includes many established effect sizes, such as odds ratio, risk difference, Cohen's d, number needed to treat, common language effect size, Glass rank-biserial correlation, and Vargha and Delaney's A, among others. Every estimate is returned with a 95% confidence interval and a $p$-value produced by a single unified process, so the interval and the $p$-value always agree and no separate significance test is needed. Output includes a plain-language interpretation of each estimate along with Methods- and Results-section text ready to paste into a manuscript.

`es4all` is the companion software package to the paper "The dominance family: interpretable effect sizes with unified estimation and inference". All of the calculations in that paper's examples were performed with this package.

## Installation

To install from GitHub:

```r
install.packages("remotes")
remotes::install_github("es4all-package/es4all", upgrade = "never")
```

## Example

`d_bin()` performs calculations for binary data. To keep this example short, we only show one effect size -- the stochastic difference (`SD`). This effect size is also known in the literature by many other names, including Glass rank-biserial correlation; net benefit; Cliff's delta; simple difference formula; Stuart tau-c; risk difference; excess risk; attributable risk; and absolute risk increase. 

```{r example, results='asis'}
set_opts(es = "SD")
d_bin(comparison_data$UTI)
```
