---
title: "Introduction to es4all"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to es4all}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The dominance family of effect sizes

For two groups, let $d_{1,2}$ be the probability that a randomly drawn observation from group 1 is greater than a randomly drawn observation from group 2, $d_{2,1}$ the probability of the reverse, and $d^0$ the probability of a tie ($d_{1,2}+d_{2,1}+d^0=1$). Any effect size built from these three comparison probabilities is dominance-based. `es4all` reports the following members of the family:

| Effect size | Abbrev. | Definition | Also known as |
|---|---|---|---|
| Stochastic difference | SD | $d_{1,2}-d_{2,1}$ | Cliff's delta, Glass rank-biserial correlation, net benefit; = risk difference (RD) for binary data |
| Adjusted stochastic difference | ASD | $(d_{1,2}-d_{2,1})/(d_{1,2}+d_{2,1})$ | Goodman and Kruskal's gamma; = Yule's Q for binary data |
| Number needed to treat | NNT | $1/(d_{1,2}-d_{2,1})$ | reciprocal of risk difference; binary data only |
| Dominance | D | $d_{1,2}$ | Common language effect size (CLES), Grissom and Kim's probability of superiority |
| Measure of stochastic superiority | MSS | $d_{1,2}+0.5\,d^0$ | Vargha and Delaney's A |
| Adjusted dominance | AD | $d_{1,2}/(d_{1,2}+d_{2,1})$ | - |
| Generalized odds ratio | GOR | $d_{1,2}/d_{2,1}$ | Agresti's generalized odds ratio, win ratio; = odds ratio (OR) for binary data; = odds for continuous data |
| Win odds | WO | $(d_{1,2}+0.5\,d^0)/(d_{2,1}+0.5\,d^0)$ | odds form of MSS |
| Standardized mean difference | SMD | $\sqrt{2} \Phi^{-1}(d_{1,2})$ | Cohen's d; continuous data only |

For discrete (binary or ordinal) data, an important property of the effect size is whether it is **equivalence dependent**. SD, MSS, WO, and NNT are equivalence dependent: their attainable range narrows as the probability of a tie grows, making them shrink toward the point of no effect. See the [binary-data example](#why-equivalence-dependence-matters) below for an illustration of why this is a desirable property.

## Functions

| Function | Data type | Estimation |
|---|---|---|
| `d_cont()` | Continuous (normal) | Maximum likelihood (MLE) |
| `d_bin()` | Binary | MLE |
| `d_ord()` | Ordinal | MLE |
| `d_one()` | Continuous, one-sample | MLE |
| `d_pair()` | Continuous, paired | MLE |
| `d_np()` | Any (continuous, binary, or ordinal) | Non-parametric (distribution-free) |

Each `d_*()` function accepts known distribution parameters, sample summary statistics, or record-level microdata as input.

## Examples

### Continuous data

```{r}
library(es4all)
```

`set_opts(es = NA)` prints all effect sizes (the default), while `set_opts(es = c("D", "SD", ...))` prints a chosen subset. 

```{r, results='asis'}
set_opts(es = c("D", "SD"))
d_cont(comparison_data$stature2)
```

### Why equivalence dependence matters

This is the myocardial infarction (MI) example that has been discussed in the effect size literature. 

```{r, results='asis'}
set_opts(es = c("SD", "GOR"))
d_bin(comparison_data$MI)
```

The equivalence-dependent SD is tiny (0.01) because almost all pairs of patients tie on the outcome ("had an MI" vs. "did not have an MI"), while the equivalence-independent GOR (1.8) describes a much larger effect among the pairs that are actually distinguishable. These descriptions answer different questions -- SD the population-wide question a treatment decision turns on, GOR the narrower question of how one-sided the comparison is among the pairs that differ.

### Ordinal data

Analysis of data from the 1948 trial of streptomycin for the treatment of tuberculosis.

```{r, results='asis'}
set_opts(es = NA) # restore the default (print all effect sizes)
d_ord(comparison_data$TB)
```

## Built-in example datasets

`es4all` comes with several datasets from published studies, used throughout its documentation and reproduced in the paper's examples:

```{r}
ret = NULL
for (nn in names(comparison_data)) {
  r1 = data.frame(Label = nn
    , About = strtrim( attr(comparison_data[[nn]], "meta")$About, 25)
    , `Group 1` = comparison_data[[nn]]$group[1]
    , `Group 2` = comparison_data[[nn]]$group[2]
    , Source = strtrim( attr(comparison_data[[nn]], "meta")$Source, 25)
    , check.names = FALSE)
  ret = rbind(ret, r1)
}
print(ret)
```

## Customizing output

```{r}
set_opts(
  conf.level = 0.95,  # confidence level of the interval
  es = NA,            # which effect sizes to print: NA (all, default) or a vector of names
  output = "auto",    # "auto", "huxtable", "gt", "kableExtra", "flextable", "screen", "Excel", "Word", "CSV"
  raw = FALSE         # if TRUE, return the raw intermediate data.frame instead of formatted output
)
show_opts()
```

`output = "auto"` selects huxtable for the console, gt for HTML, or kableExtra for PDF/LaTeX. `"Excel"`, `"Word"`, and `"CSV"` write to a file (pass `file = "..."`) and require the `openxlsx2` (for Excel) or `flextable` + `officer` (for Word) packages.
